Resolvable Ambiguity

### 2019-07-11 paper/paper.tex
Weaken abstract

%% Abstract
%% Note: \begin{abstract}...\end{abstract} environment must come
%% before \maketitle command
\begin{abstract}
A common standpoint when designing the syntax of programming languages is that
the grammar definition has to be unambiguous.  However, requiring up front unambiguous grammars
can force language designers to make more or less arbitrary choices to disambiguate
the language. In this paper, we depart from the traditional view of
unambiguous grammar design, and enable the detection of ambiguities to be delayed until
parse time, allowing the user of the language to perform the disambiguation. A
natural decision problem follows: given a language definition, can a
user always disambiguate an ambiguous program?  We introduce and
formalize this fundamental problem---called the \emph{resolvable
  ambiguity problem}---and divide it into separate static and dynamic
resolvability problems. We provide solutions to the static problem
for two restricted classes of languages and prove them sound and,
depending on the restrictions, complete. We also provide a sound
and complete solution to the dynamic problem for a much less
restricted class of languages.
%
The approach is evaluated through two separate case studies,
covering both a large existing programming language, and the
composability of domain-specific languages.
\end{abstract}

>>>>

%% Abstract
%% Note: \begin{abstract}...\end{abstract} environment must come
%% before \maketitle command
\begin{abstract}
A common standpoint when designing the syntax of programming languages is that
the grammar definition has to be unambiguous.  However, requiring up front unambiguous grammars
can force language designers to make more or less arbitrary choices to disambiguate
the language. In this paper, we depart from the traditional view of
unambiguous grammar design, and enable the detection of ambiguities to be delayed until
parse time, allowing the user of the language to perform the disambiguation. A
natural decision problem follows: given a language definition, can a
user always disambiguate an ambiguous program?  We introduce and
formalize this fundamental problem---called the \emph{resolvable
  ambiguity problem}---and divide it into separate static and dynamic
resolvability problems. We provide solutions to the static problem
for two restricted classes of languages and sketch proofs of soundness and,
depending on the restrictions, completeness. We also provide a sound
and complete solution to the dynamic problem for a much less
restricted class of languages.
%
The approach is evaluated through two separate case studies,
covering both a large existing programming language, and the
composability of domain-specific languages.
\end{abstract}
